Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-10 ms.tex
Fixing references

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the galactic nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{Heckman&Best14}).  SMBH growth
in the low redshift universe is dominated by low mass black holes,
$M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a
fact which is often attributed to the general trend of `cosmic
down-sizing' resulting from hierarchical structure growth (e.g,
\citealt{Gallo+08}).  However, the physical processes involved in
growing low mass blacks may be distinct from those operating at higher
SMBH masses and accretion rates in quasars.  A key open question is
whether SMBHs grow primarily by accreting gas fed in directly from
galactic or extra-galactic scales, or whether significant growth
results also from internal, secular processes such as local stellar
mass loss in the nuclear region.

@@

The gas density in galactic nuclei also influences the emission from
stellar tidal disruption events (TDEs), such as the high energy
transient {\it Swift} J1644+57 (\citealt{Levan+11};
\citealt{Bloom+11}, \citealt{Burrows+11}; \citealt{Zauderer+11}).
This transient event was powered by an impulsive relativistic jet,
which produced synchrotron radio emission as the jet material was
decelerated by shock interaction with the CNM of the previously
quiescent SMBH (\citealt{Giannios&Metzger11}; \citealt{Zauderer+11}).
Detailed modeling of J1644+57 showed that the CNM density was much
lower than that measured surrounding Sgr A$^{\star}$
(\citealt{Metzger+12}; \citealt{Berger+12}).  However, a TDE jet
injected into a denser CNM would be decelerated more rapidly,
producing substantially different radio emission than in J1644+57.
Variations in the properties of the CNM could help explain why most
TDEs appear to be radio quiet (e.g.~\citealt{Bower+13};
\citealt{VanVelzen+13}).

@@

A more readily addressable test is whether a steady-state accretion
picture is capable of explaining the large scatter of typically 2-3
orders of magnitude in the observed distribution of $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter is expected to result from at
least two effects: (1) the strong sensitivity of the accretion rate to
the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) in
combination with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (~\ref{fig:NickPlot2}) and
(2) angular momentum of the infalling gas and consequentiallly the
size of small scale outflows may vary from galaxy to galaxy at fixed
$\Mbh$. The combination of (1) and (2) is capable of explaining much
of the observed sample, especially considering that scatter in $L_{\rm
  X}$ due to variations in $\dot{M}$ is exacerbated by the expectation
that $\epsilon_{X} \propto \dot{M}$ for radiatively inefficient
flows. At a wind heating parameter there is an order of magnitude
difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$
and $M_{\bullet}$ (Fig.~\ref{fig:bh_growth}) which could also lead to
scatter of $\dot{M}$ due to the variation in the Nuker stellar density
profile $\Gamma$ at fixed $M_{\bullet}$. However, \citet{Pellegrini10}
find that the scatter in $L_X$ is similar if one just looks at
core/cusp galaxies separately. There could also be scatter introduced
in using scaling relationships ($\Mbh-\sigma$) and ($\Mbh-L_{\rm
  bulge}$) to measure $\Mbh$ as \citet{Miller+2015} do.

@@

Our results also have implications for the environments encountered by
relativistic jets from TDEs.  For low-mass SMBHs $M_{\bullet} \lesssim
10^{7}M_{\odot}$, characterizing those responsible for {\it Swift}
J1644+57 (\citealt{Bloom+2011}), we predict gas densities of a $n
\sim 0.3-3$ cm$^{-3}$ on radial scales of 0.3 pc for wind heating
rates within the physically expected range $v_{\rm w} \sim 500-1000$
km s$^{-1}$ (Fig.~\ref{fig:profiles}).  These values are within the
range estimated by modeling the radioa afterglow of {\it Swift}
J1644+57 (\citealt{Metzger+12}, \citealt{Berger+12}).  In fact, the
flattening of the density profile inferred by \citet{Berger+12} on
radii scales $r \gtrsim 0.4$ pc (their Fig.~6) looks similar to that
seen in our calculations of the gas density profile for core galaxies
(Fig.~\ref{fig:profiles}, solid line).  The CNM density of the host of
{\it Swift} J1644+57 was a factor of 30 times lower than that inferred
by \citet{Quataert:2004a} for SgrA*, a fact we attribute to...{\bf
  BDM: whats going on?}  The implications of our models for CNM
profiles for the diversity of radio emission from TDE jets will be
explored further in future work.

>>>>

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the galactic nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  SMBH growth
in the low redshift universe is dominated by low mass black holes,
$M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a
fact which is often attributed to the general trend of `cosmic
down-sizing' resulting from hierarchical structure growth (e.g,
\citealt{Gallo+08}).  However, the physical processes involved in
growing low mass blacks may be distinct from those operating at higher
SMBH masses and accretion rates in quasars.  A key open question is
whether SMBHs grow primarily by accreting gas fed in directly from
galactic or extra-galactic scales, or whether significant growth
results also from internal, secular processes such as local stellar
mass loss in the nuclear region.

@@

The gas density in galactic nuclei also influences the emission from
stellar tidal disruption events (TDEs), such as the high energy
transient {\it Swift} J1644+57 (\citealt{Levan+11};
\citealt{Bloom+11}, \citealt{Burrows+11}; \citealt{Zauderer+11}).
This transient event was powered by an impulsive relativistic jet,
which produced synchrotron radio emission as the jet material was
decelerated by shock interaction with the CNM of the previously
quiescent SMBH (\citealt{Giannios&Metzger11}; \citealt{Zauderer+11}).
Detailed modeling of J1644+57 showed that the CNM density was much
lower than that measured surrounding Sgr A$^{\star}$
(\citealt{Metzger+12}; \citealt{BergerZauderer+:2012a}).  However, a TDE jet
injected into a denser CNM would be decelerated more rapidly,
producing substantially different radio emission than in J1644+57.
Variations in the properties of the CNM could help explain why most
TDEs appear to be radio quiet (e.g.~\citealt{Bower+13};
\citealt{VanVelzen+13}).

@@

A more readily addressable test is whether a steady-state accretion
picture is capable of explaining the large scatter of typically 2-3
orders of magnitude in the observed distribution of $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter is expected to result from at
least two effects: (1) the strong sensitivity of the accretion rate to
the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) in
combination with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (~\ref{fig:NickPlot2}) and
(2) angular momentum of the infalling gas and consequentiallly the
size of small scale outflows may vary from galaxy to galaxy at fixed
$\Mbh$. The combination of (1) and (2) is capable of explaining much
of the observed sample, especially considering that scatter in $L_{\rm
  X}$ due to variations in $\dot{M}$ is exacerbated by the expectation
that $\epsilon_{X} \propto \dot{M}$ for radiatively inefficient
flows. At a wind heating parameter there is an order of magnitude
difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$
and $M_{\bullet}$ (Fig.~\ref{fig:bh_growth}) which could also lead to
scatter of $\dot{M}$ due to the variation in the Nuker stellar density
profile $\Gamma$ at fixed $M_{\bullet}$. However, \citet{Pellegrini10}
find that the scatter in $L_X$ is similar if one just looks at
core/cusp galaxies separately. There could also be scatter introduced
in using scaling relationships ($\Mbh-\sigma$) and ($\Mbh-L_{\rm
  bulge}$) to measure $\Mbh$ as \citet{Miller+15} do.

@@

Our results also have implications for the environments encountered by
relativistic jets from TDEs.  For low-mass SMBHs $M_{\bullet} \lesssim
10^{7}M_{\odot}$, characterizing those responsible for {\it Swift}
J1644+57 (\citealt{Bloom+11}), we predict gas densities of a $n
\sim 0.3-3$ cm$^{-3}$ on radial scales of 0.3 pc for wind heating
rates within the physically expected range $v_{\rm w} \sim 500-1000$
km s$^{-1}$ (Fig.~\ref{fig:profiles}).  These values are within the
range estimated by modeling the radioa afterglow of {\it Swift}
J1644+57 (\citealt{Metzger+12}, \citealt{BergerZauderer+:2012a}).  In fact, the
flattening of the density profile inferred by \citet{BergerZauderer+:2012a} on
radii scales $r \gtrsim 0.4$ pc (their Fig.~6) looks similar to that
seen in our calculations of the gas density profile for core galaxies
(Fig.~\ref{fig:profiles}, solid line).  The CNM density of the host of
{\it Swift} J1644+57 was a factor of 30 times lower than that inferred
by \citet{Quataert:2004a} for SgrA*, a fact we attribute to...{\bf
  BDM: whats going on?}  The implications of our models for CNM
profiles for the diversity of radio emission from TDE jets will be
explored further in future work.

### 2015-03-10 windheat.tex
Fixing references

Figure \ref{NickPlot2} shows how the wind heating varies as star formation histories deviate from their halo-averaged values.  In particular, we show
\begin{align} 
  \dot{\mathcal{M}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{\mathcal{E}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1,\\
  \mathcal{V}_w^2(t) &=2 \dot{\mathcal{E}}(t)/\dot{\mathcal{M}}(t).
\end{align}
In these equations, 
\begin{equation}
\mathcal{S}(t) = S(t) \times \left(\frac{2}{\pi}(1-\zeta)\arctan(t/t_{\rm scale}) + \zeta \right).
\end{equation}
This function convolves the recent ($z \approx 0$) halo-averaged star formation history with local variation to give a more pessimistic estimate for the value of $\tilde{v}_{\rm w}$.  In particular, replacing $S(t)$ with $\mathcal{S}(t)$ reduces the recent star formation rate to a fraction $\epsilon$ of its halo-averaged value, and does so for a characteristic time $t_{\rm scale}$ into the past.  As we can see in Fig. \ref{NickPlot2}, this dramatically lowers the effective wind speed when both $t_{\rm scale} \gtrsim 10^7 ~{\rm yr}$ and $\epsilon \lesssim 0.1$, but otherwise has too modest of an effect to change the thermal stability properties of the flow (although the location of $r_{\rm s}$ and the value of $\dot{M}$ may change significantly).

>>>>

Figure \ref{fig:NickPlot2} shows how the wind heating varies as star formation histories deviate from their halo-averaged values.  In particular, we show
\begin{align} 
  \dot{\mathcal{M}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{\mathcal{E}}(t) &= \int_0^t \mathcal{S}(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1,\\
  \mathcal{V}_w^2(t) &=2 \dot{\mathcal{E}}(t)/\dot{\mathcal{M}}(t).
\end{align}
In these equations, 
\begin{equation}
\mathcal{S}(t) = S(t) \times \left(\frac{2}{\pi}(1-\zeta)\arctan(t/t_{\rm scale}) + \zeta \right).
\end{equation}
This function convolves the recent ($z \approx 0$) halo-averaged star formation history with local variation to give a more pessimistic estimate for the value of $\tilde{v}_{\rm w}$.  In particular, replacing $S(t)$ with $\mathcal{S}(t)$ reduces the recent star formation rate to a fraction $\epsilon$ of its halo-averaged value, and does so for a characteristic time $t_{\rm scale}$ into the past.  As we can see in Fig. \ref{fig:NickPlot2}, this dramatically lowers the effective wind speed when both $t_{\rm scale} \gtrsim 10^7 ~{\rm yr}$ and $\epsilon \lesssim 0.1$, but otherwise has too modest of an effect to change the thermal stability properties of the flow (although the location of $r_{\rm s}$ and the value of $\dot{M}$ may change significantly).
